\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-19 notes/2loopNLSM.tex
Alex partial readthrough.

  We make progress towards understanding how color-dual representations must be constructed at the multi-loop level. First
  identify a cubic theory, \emph{semi-abelian} Yang--Mills, that unifies many of the color-dual theories constructed in the literature, and provides a construction of color-dual numerators at one-loop directly from Feynman rules. Joint
  consideration of these theories highlights that the lifting of
  one-loop color-kinematics dual numerators to two loops may be just
  as challenging as the step from tree level to one loop.  As a first
  step in understanding the two-loop challenge, we construct
  color-dual two-loop four-point non-linear sigma model (NLSM)
  numerators using generalized unitarity methods.  The double-copy of
  the constructed NLSM numerators with themselves, $\mathcal{N}=0$
  Yang--Mills, and $\mathcal{N}=4$ Yang--Mills correctly reproduce the
  known physical unitarity cuts of special Gallileons, Born-Infeld
  theory, and Dirac-Born-Infeld-Volkov-Akulov theory, respectively.
  These NLSM numerators are in explicit tension with generalizations
  of the one-loop NLSM numerators. \ace{I realize that we all know
    this, but does/will a statement to this effect actually make it
    into the draft?}  Additionally, we revisit the two-loop four-point
  Yang--Mills problem in order to identify the exact point of
  inconsistency in a local color-dual construction.

@@

However, despite the tremendous success in leveraging this duality to compute gauge and gravity observables to high orders in perturbation theory \cite{FiveLoopN4,GeneralizedDoubleCopyFiveLoops,Bern:2018jmv,Carrasco:2021otn}, color-kinematics remains a conjecture at loop level. Indeed, there are many examples in the literature where identifying color-dual representations beyond one-loop has posed a formidable challenge \cite{Mogull:2015adi,Johansson:2017bfl,KalinN2TwoLoop}. In this work, we use the nonlinear sigma model (NLSM) and Yang-Mills, two theories proven to permit color-dual representations at tree-level \cite{Feng:2010my,Cachazo:2014xea}, as case studies for advancing our understanding of the kinematic algebra to the multi-loop level.

@@

Faced with this obstruction, we study the multi-loop sector of both NLSM and Yang-Mills in \cref{2loopBoot} using an ansatz based color-dual bootstrap. In contrast to much of the available literature on construction of color-dual integrands at two-loop \cite{Bern:2013yya,Mogull:2015adi,Johansson:2017bfl}, our results are completely agnostic to the spacetime dimension; all the polarizations and momenta appearing in our construction will remain formally $D$-dimensional. This makes our methods and results particularly well suited for algorithmically extracting rational terms from the $D$-dependence of dimensionally regulated loop momenta. We are able to achieve this by employing a custom solver, \texttt{FiniteFieldSolve}, which will soon be made public.

With this solver, we are able to compute a two-loop integrand for NLSM that globally manifests the duality off-shell. This represents a new benchmark for color-dual representations in non-supersymmetric gauge theories. However in contrast, we find surprisingly that Yang-Mills does not permit a globally color-dual representation, even when considering the most general polynomial ansatz of Lorentz covariant kinematics. Concretely, we find a sharp failure point between the ``bowtie'' cut, and the following jacobi triple:
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
We comment on the implications of this finding in \cref{conclusions}, and discuss alternative approaches where one must relax field theoretic locality in order to accommodate the construction of globally color-dual multi-loop integrands.

@@

I think this belongs in another section...
Why DC is cool:
1) Reduction in computational complexity.
Off-shell high spin is terrible.
Gravity has infinitely many very complex Feynman rules as a result of gauge redundancy but YM only has two relatively simple Feynman rules.
2) DC points to deep structure hidden in Lagrangian formulation.

Color-kinematics duality is a statement about the off-shell graphical simplicity that manifest redundancy in the on-shell observables of gauge and gravity theories. Ignoring factors of $i$ and the coupling constant, a generic $m$-point $L$-loop amplitude in a color-dual theory -- such as NLSM, YM, or bi-adjoint scalar theory -- can be written as
\be
\mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
\ee
Here the spacetime dimension is $D$.
Even though the underlying theory may have quartic (or higher multiplicity) vertices, it is possible to multiply and divide by propagators so that the sum runs over only purely cubic graphs $\Gamma$.
Overcounting is removed by the internal symmetry factor $S_\Gamma$, which tracks the number of graph symmetries with external legs held fixed.
The color factor $c_\Gamma$ is the string of structure constants $f^{abc}$ associated with the graph $\Gamma$,\footnote{This paper is focused on theories with particles that transform in the adjoint.  However, the double copy can be applied to a far more diverse set of theories with particles that transform in the fundamental or more diverse representations, see for example \cite{Johansson:2017srf, Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2023vjg, Carrasco:2020ywq}.}
the denominator $d_\Gamma$ is the product of propagators associated with the graph $\Gamma$, and the numerator $n_\Gamma$ is all of the remaining kinematic factors including dot products of momenta and polarization vectors.

The color factors $c_\Gamma$ are not independent and are related to one another via a set of Jacobi identities that look schematically like
\be
c_i+c_j+c_k=0,
\ee
where $i$, $j$, and $k$ are three graphs.
Example Jacobi relations are presented in detail in \cref{sec:bootstrap}.
Color-kinematics duality states that there exists a way to rearrange factors between the various kinematic numerators $n_\Gamma$ such that the numerators obey all of the same Jacobi identities as the color factors
\be
n_i +n_j +n_k=0.
\ee
If color-kinematics duality can be achieved at the off-shell level, then the kinematic numerators should satisfy an algebra just like the color factors.
In theories where the kinematic algebra is known, it is often the algebra of volume preserving diffeomorphisms \cite{Monteiro2011pc, Ben-Shahar:2021zww, Cheung:2020djz, Cheung:2021zvb, Cheung:2022mix}.
As a simple example, color-dual numerators for 4pt pion scattering at tree level are
\begin{align}
&n_s = s(t-u)\\
&n_t = t(u-s)\\
&n_u = u(s-t)
\end{align}
where the coupling has been normalized away.
These numerators sum to zero by explicit calculation.
Note that the $s$-channel numerator contains an explicit factor of $s$ that cancels with the propagator and the same goes for the $t$- and $u$-channels so that the net result is a local function, just as one would expect for the pion tree amplitude.

Once color-dual numerators $n_\Gamma$ have been found, they can be used to replace the color factors ($c_\Gamma \to n_\Gamma$) in a \emph{different} amplitude to produce the corresponding amplitude in the double-copy theory
\be\label{DCconstruction}
\tilde{\mathcal{A}}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma \tilde{n}_\Gamma}{d_\Gamma} ~~ \xrightarrow{(c_\Gamma \to n_\Gamma)} ~~ \mathcal{M}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{n_\Gamma \tilde{n}_\Gamma}{d_\Gamma}.
\ee
Importantly, the $\tilde{n}_\Gamma$ do not need to respect color-kinematics duality.
The prototypical example of the double-copy is that YM double copied with YM results in gravity where the gauge invariance of each separate gluon produces the diffeomorphism invariance of the graviton.
The double copy can be proven at tree level using the KLT relations but there is significant evidence that the double copy persists to all loop orders \cite{KLT, KiermaierTalk} \citepls{Any loop papers we want?}.
In addition to pointing at deep structure hidden in the Lagrangian for gravity, the double copy is immensely useful at a utilitarian level since it is much more efficient at producing $\mathcal{M}_m^L$ than traditional methods like Feynman rules.

Many of the 1-loop results presented in \cref{sec:cubic} are informed by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ in YM, NLSM, or bi-adjoint scalar theory can be decomposed in terms of the trace basis
\be
\mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
\ee
where $T^a$ are generators of the gauge group and, because of the cyclicity of the trace, one of the legs is held fixed so that the sum runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial amplitudes $A_n[...]$.
Any tree-level color factor can be converted into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder color factors
\be
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\ee
associated with the graph
\be
\halfLadder
\ee
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\be
\mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
\ee
with legs $1$ and $n$ fixed so that the sum only contains $(n-2)!$ terms, indicating that there are additional relations -- the Kleiss-Kuijf (KK) relations -- amongst the partial amplitudes \cite{Kleiss:1988ne}.
The KK relations are generic to any theory with purely adjoint particles.
Color-kinematics duality implies that the KK basis is overcomplete and can be reduced to a basis of $(n-3)!$ amplitudes via the fundamental BCJ identities \cite{BCJ,Feng:2010my}
\be
\sum \limits_{i=2}^{n-1} k_1 \cdot (k_2+...+k_i) A_n[2,...,i,1,i+1,...,n] =0.
\ee
In terms of constructing explicit color-dual solutions at tree level, it is enough to specify the numerator of the half ladder because any diagram can be reduced to a half ladder through successive applications of the Jacobi identity.
At 1-loop, every diagram can be related to one topology as well, in this case the $n$-gon, but the pattern does not persist to 2-loops.

\subsection{Color-dual Lagrangians}\label{offShellCK}
While BCJ relations is an on-shell statement about the color-dual nature of scattering amplitudes, one might aspire to hard code color-kinematics duality by realizing an off-shell Lagrangian description of the theory. The double-copy construction of \ref{DCconstruction} would suggest a manifestly cubic Lagrangian of the form:
\be
\mathcal{L} = P_V^{ab}P_W^{ij}\mathcal{O}^a_i \Box \mathcal{O}^b_j +  V^{abc} W^{ijk}\mathcal{O}^a_i \mathcal{O}^b_j \mathcal{O}^c_k
\ee
where $V^{abc}$ and $W^{ijk}$ are cubic Feynman rules mixing field operators $\mathcal{O}^{a}_i$ indexed by quantum numbers $a$ and $i$. We have introduced $P_V^{ab}$ and $P_W^{ij}$ projection operators to encode non-local kinetic structure that could in principle participate in the construction. Should the off-shell Feynman rules of this cubic theory be antisymmetric and satisfy Jacobi identities on all internal edges,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0\,,
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0\,,
\end{align}
where 
\be
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv V^{a_1a_2b_1}V^{b_1a_3b_2}\cdots V^{b_{n-3} a_{n-2}a_n}\,.
\ee
This is a sufficient condition of the amplitudes to obey BCJ relations. Off-shell description of color-dual theories are rare, and only a few examples are available in the literature. Typically when one goes about construction color-dual theories directly from a set of cubic interactions, higher point Jacobi and antisymmetry constraints require introducing additional operators. See refs \cite{Ben-Shahar:2022ixa} for some recent examples. Indeed, as we'll show in \cref{sec:cubic}, Yang-Mills requires introducing a cubic 2-form interaction in order to manifest the kinematic algebra just at four-point. Absent information about the kinematic algebra, one can make progress in the construction of loop-level color-dual amplitudes by employing a color-dual bootstrap. 
\subsection{Color-dual Bootstrap}
\label{sec:bootstrap}

Aside from a few special theories that permit an off-shell Lagrangian description \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb, Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww, Ben-Shahar:2021doh}, color-dual integrands are constructed from an ansatz.
The four conditions we impose on an ansatz are summarized here and elaborated on below.\footnote{The four ansatz conditions are similar to those of Ref.~\cite{Bern:2015ooa} except that the authors demand that the numerators have the same loop power counting as Feynman gauge Feynman rules.  This is an additional assumption that is most emphatically not made here.}
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials in momenta and
  polarizations, \eg the diagrams only have poles given by the
  propagators of the diagrams.
  \item  \textbf{Color-kinematics duality}: The numerators obey the same Jacobi identities as the color factors.
This implicitly requires that the integrand is expressed in terms of purely cubic diagrams.
\item  \textbf{Graph symmetries}: The diagram numerators are only a function of the diagram's topology and labeling.  Furthermore, the diagram numerators are invariant under the automorphisms of the diagrams \emph{including signs that compensate for color-factor sign changes}.
\item  \textbf{On-shell Unitarity }: The cuts of the ansatz must reproduce the physical unitarity cuts of the theory when internal momenta are taken on-shell.
\end{enumerate}

In order to clarify general statements, the following subsections include examples from a 4pt 2-loop integrand for some color-dual theory like YM or NLSM.
This is the simplest integrand that demonstrates the full machinery of the color-dual bootstrap.
All tree and 1-loop processes are generated by single basis diagrams (the half ladder and $n$-gon respectively) so they lack examples of the ``boundary'' Jacobi relations to be described shortly.
Choosing the 4pt 2-loop integrand also has the advantage that this process appears prominently in this paper.

@@

The kinematic numerators must be $D$-dimensional, Lorentz invariant, local (polynomial) functions of the graph kinematics with the correct power counting and external states.
Insisting on a local numerator is partly a matter of convenience as the ansatz for a rational function would quickly grow out of control without a guiding principle for what terms to include.
Locality is also physically motivated since Feynman rules typically produce polynomial numerators.
While these assumptions are well motivated from a physical and complexity standpoint, the literature has featured numerators that violate locality, modify naive power counting through the inclusion of higher spin modes, and break manifest Lorentz invariance \cite{Square, WeinzierlBCJLagrangian, Mogull:2015adi, FivePointN4BCJ, Johansson:2017bfl, Ben-Shahar:2022ixa, Cheung:2016prv, Cheung:2021zvb}.
These developments suggest that relaxing locality is a viable alternative when the condition proves too stringent for an ansatz.
We will comment more on locality in \sect{}.

@@

\paragraph{Color-kinematics duality}

Color-kinematics duality is a statement about color factors, kinematic numerators, and the accompanying Jacobi relations.
Adjoint-type color factors are associated with cubic graphs, so satisfying color-kinematics duality implicitly requires expressing the integrand in terms of such graphs. As mentioned earlier, this is possible to do even for a theory with quartic (or higher multiplicity) vertices by multiplying and dividing by propagators.
In the case of the 4pt 2-loop example promised above, there are 14 cubic graph topologies (shown in \cref{fig:MaxCuts}) ignoring tadpoles and bubble on external leg (BEL) graphs.

>>>>

  We make progress towards understanding how color-dual
  representations must be constructed at the multi-loop level. First
  we identify a cubic theory, \emph{semi-abelian} Yang--Mills, that
  unifies many of the color-dual theories constructed in the
  literature, and provides a construction of color-dual numerators at
  one-loop directly from Feynman rules. Joint consideration of these
  theories highlights that the lifting of one-loop color-kinematics
  dual numerators to two loops may be just as challenging as the step
  from tree level to one loop.  As a first step in understanding the
  two-loop challenge, we construct color-dual two-loop four-point
  non-linear sigma model (NLSM) numerators using generalized unitarity
  methods.  The double-copy of the constructed NLSM numerators with
  themselves, $\mathcal{N}=0$ Yang--Mills, and $\mathcal{N}=4$
  Yang--Mills correctly reproduce the known physical unitarity cuts of
  special Gallileons, Born-Infeld theory, and
  Dirac-Born-Infeld-Volkov-Akulov theory, respectively.  These NLSM
  numerators are in explicit tension with generalizations of the
  one-loop NLSM numerators. \ace{I realize that we all know this, but
    does/will a statement to this effect actually make it into the
    draft?}  Additionally, we revisit the two-loop four-point
  Yang--Mills problem in order to identify the exact point of
  inconsistency in a local color-dual construction.

@@

However, despite the tremendous success in leveraging this duality to
compute gauge and gravity observables to high orders in perturbation
theory \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv,
  Carrasco:2021otn}, color-kinematics remains a conjecture at loop
level. Indeed, there are many examples in the literature where
identifying color-dual representations beyond one-loop has posed a
formidable challenge
\cite{Mogull:2015adi,Johansson:2017bfl,KalinN2TwoLoop} \ace{BDN might
  fit better here rather than the 2loop SQCD}. In this work, we use
the nonlinear sigma model (NLSM) and Yang-Mills, two theories proven
to permit color-dual representations at tree-level
\cite{Feng:2010my,Cachazo:2014xea}, as case studies for advancing our
understanding of the kinematic algebra to the multi-loop level.

@@

Faced with this obstruction, we study the multi-loop sector of both
NLSM and Yang-Mills in \cref{2loopBoot} using an ansatz based
color-dual bootstrap. In contrast to much of the available literature
on construction of color-dual integrands at two-loop
\cite{Bern:2013yya,Mogull:2015adi,Johansson:2017bfl}, our results are
completely agnostic to the spacetime dimension; all the polarizations
and momenta appearing in our construction will remain formally
$D$-dimensional. This makes our methods and results particularly well
suited for algorithmically extracting rational terms from the
$D$-dependence of dimensionally regulated loop momenta. We are able to
achieve this by employing a custom solver, \texttt{FiniteFieldSolve},
which will soon be made public. \ace{This either needs an extra
  context sentence, or moved out of the intro.}

With this solver, we are able to compute a two-loop integrand for NLSM
that globally manifests the duality off-shell. This represents a new
benchmark for color-dual representations in non-supersymmetric gauge
theories. However in contrast, we find surprisingly that Yang-Mills
does not permit a globally color-dual representation, \ace{Maybe tone
  down surprise and ref BDN} even when considering the most general
polynomial ansatz of Lorentz covariant kinematics. Concretely, we find
a sharp failure point between the ``bowtie'' cut, and the following
jacobi triple:
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
We comment on the implications of this finding in \cref{conclusions},
and discuss alternative approaches where one must relax field
theoretic locality in order to accommodate the construction of
globally color-dual multi-loop integrands.

@@

%I think this belongs in another section...
%Why DC is cool:
%1) Reduction in computational complexity.
%Off-shell high spin is terrible.
%Gravity has infinitely many very complex Feynman rules as a result of gauge redundancy but YM only has two relatively simple Feynman rules.
%2) DC points to deep structure hidden in Lagrangian formulation.

Color-kinematics duality is a statement about the off-shell graphical
simplicity that manifest redundancy in the on-shell observables of
gauge and gravity theories. Ignoring factors of $i$ and the coupling
constant, a generic $m$-point $L$-loop amplitude in a color-dual
theory -- such as NLSM, YM, or bi-adjoint scalar
theory\ace{technically this is true as written for any theory.  In
  fact we slip that type of disclaimer in quite a lot throughout the
  rest of the section} -- can be written as
\begin{equation}
\mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
\end{equation}
Here the spacetime dimension is $D$.  Even though the underlying
theory may have quartic (or higher multiplicity) vertices, it is
possible to multiply and divide by propagators so that the sum runs
over only purely cubic graphs $\Gamma$.  Overcounting is removed by
the internal symmetry factor $S_\Gamma$, which tracks the number of
graph symmetries with external legs held fixed.  The color factor
$c_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu,
    Carrasco:2023vjg, Carrasco:2020ywq}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $n_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors.

The color factors $c_\Gamma$ are not independent and are related to
one another via a set of Jacobi identities that look schematically
like
\begin{equation}
c_i+c_j+c_k=0,
\end{equation}
where $i$, $j$, and $k$ are three graphs.  Example Jacobi relations
are presented in detail in \cref{sec:bootstrap}.  Color-kinematics
duality states that there exists a way to rearrange factors between
the various kinematic numerators $n_\Gamma$ such that the numerators
obey all of the same Jacobi identities as the color factors
\begin{equation}
n_i +n_j +n_k=0.
\end{equation}
If color-kinematics duality can be achieved at the off-shell level,
then the kinematic numerators should satisfy an algebra just like the
color factors.  In theories where the kinematic algebra is known, it
is often the algebra of volume preserving diffeomorphisms
\cite{Monteiro2011pc, Ben-Shahar:2021zww, Cheung:2020djz,
  Cheung:2021zvb, Cheung:2022mix}.  As a simple example, color-dual
numerators for 4pt pion scattering at tree level are
\begin{align}
&n_s = s(t-u)\\
&n_t = t(u-s)\\
&n_u = u(s-t)
\end{align}
where the coupling has been normalized away.  These numerators sum to
zero by explicit calculation.  Note that the $s$-channel numerator
contains an explicit factor of $s$ that cancels with the propagator
and the same goes for the $t$- and $u$-channels so that the net result
is a local function, just as one would expect for the pion tree
amplitude.

Once color-dual numerators $n_\Gamma$ have been found, they can be
used to replace the color factors ($c_\Gamma \to n_\Gamma$) in a
\emph{different} amplitude to produce the corresponding amplitude in
the double-copy theory \be\label{DCconstruction}
\tilde{\mathcal{A}}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma
  \tilde{n}_\Gamma}{d_\Gamma} ~~ \xrightarrow{(c_\Gamma \to n_\Gamma)}
~~ \mathcal{M}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{n_\Gamma
  \tilde{n}_\Gamma}{d_\Gamma}.
\end{equation}
Importantly, the $\tilde{n}_\Gamma$ do not need to respect
color-kinematics duality.  The prototypical example of the double-copy
is that YM double copied with YM results in gravity where the gauge
invariance of each separate gluon produces the diffeomorphism
invariance of the graviton.  The double copy can be proven at tree
level using the KLT relations but there is significant evidence that
the double copy persists to all loop orders \cite{KLT, KiermaierTalk}
\citepls{Any loop papers we want?}.  In addition to pointing at deep
structure hidden in the Lagrangian for gravity, the double copy is
immensely useful at a utilitarian level since it is much more
efficient at producing $\mathcal{M}_m^L$ than traditional methods like
Feynman rules.

Many of the 1-loop results presented in \cref{sec:cubic} are informed
by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ in YM, NLSM, or
bi-adjoint scalar theory can be decomposed in terms of the trace basis
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
  \label{eq:trace-basis}
\end{equation}
where $T^a$ are generators of the gauge group and, because of the
cyclicity of the trace, one of the legs is held fixed so that the sum
runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial
amplitudes $A_n[...]$.  Any tree-level color factor can be converted
into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder
color factors
\begin{equation}
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\end{equation}
associated with the graph
\begin{equation}
\halfLadder
\end{equation}
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
  \label{eq:ddm-basis}
\end{equation}
with legs $1$ and $n$ fixed so that the sum only contains $(n-2)!$
terms, indicating that there are additional relations -- the
Kleiss-Kuijf (KK) relations \cite{Kleiss:1988ne} -- amongst the
partial amplitudes appearing in the right-hand side of
\cref{eq:trace-basis}.  The KK relations are generic to any theory
with purely adjoint particles.  Color-kinematics duality further
implies that the KK basis is overcomplete and can be reduced to a
basis of $(n-3)!$ amplitudes via the fundamental BCJ identities
\cite{BCJ,Feng:2010my}
\begin{equation}
\sum \limits_{i=2}^{n-1} k_1 \cdot (k_2+...+k_i) A_n[2,...,i,1,i+1,...,n] =0.
\end{equation}
In terms of constructing explicit color-dual solutions at tree level,
it is enough to specify the numerator of the half ladder because any
diagram can be reduced to a half ladder through successive
applications of the Jacobi identity.  At 1-loop, every diagram can be
related to one topology as well, in this case the $n$-gon, but the
pattern does not persist to 2-loops.

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to hard code
color-kinematics duality by realizing an off-shell Lagrangian
description of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form:
\begin{equation}
\mathcal{L} = P_V^{ab}P_W^{ij}\mathcal{O}^a_i \Box \mathcal{O}^b_j +  V^{abc} W^{ijk}\mathcal{O}^a_i \mathcal{O}^b_j \mathcal{O}^c_k
\end{equation}
where $V^{abc}$ and $W^{ijk}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_i$ indexed by quantum numbers $a$ and
$i$. We have introduced $P_V^{ab}$ and $P_W^{ij}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction. Should the off-shell Feynman rules of
this cubic theory be antisymmetric and satisfy Jacobi identities on
all internal edges,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0\,,
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0\,,
\end{align}
where 
\begin{equation}
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv V^{a_1a_2b_1}V^{b_1a_3b_2}\cdots V^{b_{n-3} a_{n-2}a_n}\,.
\end{equation}
This is a sufficient condition of the amplitudes to obey BCJ
relations. \ace{Reword previous 2 sentences} Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}\ace{Are these a reasonable
  collection?  Just grabbed the ones below}. Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional
operators. See refs \cite{Ben-Shahar:2022ixa} for some recent
examples. Indeed, as we'll show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic 2-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

\subsection{Color-dual Bootstrap}
\label{sec:bootstrap}

Aside from a few special theories that permit an off-shell Lagrangian
description \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}, color-dual integrands are constructed from an
ansatz.  The four conditions we impose on an ansatz are summarized
here and elaborated on below.\footnote{The four ansatz conditions are
  similar to those of Ref.~\cite{Bern:2015ooa} except that there
  authors demand that the numerators have the same loop power counting
  as Feynman gauge Feynman rules.  This is an additional assumption
  that is most emphatically not made here.}
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials
  in momenta and polarizations, \eg the diagrams only have poles given
  by the propagators of the diagrams.
\item \textbf{Color-kinematics duality}: The numerators obey the same
  Jacobi identities as the color factors.  This implicitly requires
  that the integrand is expressed in terms of purely cubic diagrams.
\item \textbf{Graph symmetries}: The diagram numerators are only a
  function of the diagram's topology and labeling.  Furthermore, the
  diagram numerators are invariant under the automorphisms of the
  diagrams \emph{including signs that compensate for color-factor sign
    changes}.
\item \textbf{On-shell Unitarity }: The cuts of the ansatz must
  reproduce the physical unitarity cuts of the theory when internal
  momenta are taken on-shell.
\end{enumerate}

In order to clarify general statements, the following subsections
include examples from a 4-point 2-loop integrand for some generic
color-dual theory like YM or NLSM.  This is the simplest case that
demonstrates the full machinery of the color-dual bootstrap.  All tree
and 1-loop processes are generated by single basis diagrams (the half
ladder and $n$-gon respectively) so they lack examples of the
``boundary'' Jacobi relations to be described shortly \ace{This is not
  strictly true.  Boundary Jacobi occur in 1L when demanding that BEL
  and/or tadpole are 0.  However, I appreciate that logic that 2L is
  the first time where an A+A+A=0 Jacobi occurs, rather than A+A=0.}.
Choosing the 4-point 2-loop integrand also has the advantage that this
process appears prominently in this paper.

@@

The kinematic numerators must be $D$-dimensional, Lorentz invariant,
local (polynomial) functions of the graph kinematics with the correct
power counting and external states.  Insisting on a local numerator is
partly a matter of convenience as the ansatz for a rational function
would quickly grow out of control without a guiding principle for what
terms to include.  Locality is also physically motivated since Feynman
rules typically produce polynomial numerators.  While these
assumptions are well motivated from a physical and complexity
standpoint, the literature has featured numerators that violate
locality, modify naive power counting through the inclusion of higher
spin modes, and break manifest Lorentz invariance \cite{Square,
  WeinzierlBCJLagrangian, Mogull:2015adi, FivePointN4BCJ,
  Johansson:2017bfl, Ben-Shahar:2022ixa, Cheung:2016prv,
  Cheung:2021zvb}.  These developments suggest that relaxing locality
is a viable alternative when the condition proves too stringent for an
ansatz.  We will comment more on locality in \cref{conclusions}.

@@

\paragraph{Color-kinematics duality}
\begin{figure}
\centering
{\MCutC, \MCutD, \sumoCut, \MCutJ,  
 \\
  \doubleBoxCut,  \crossBoxCut,\pentaTriangleCut,  \MCutF, 
\\
\MCutI,  \MCutE, \MCutK, 
\\
 \MCutL, \MCutG,  \MCutH,
}
\caption{The 14 non-pathological 4-point 2-loop cubic diagrams,
  including all dangling tree diagrams.} \label{fig:MaxCuts}
  \end{figure}

Color-kinematics duality is a statement about color factors, kinematic
numerators, and the accompanying Jacobi relations.  Adjoint-type color
factors are associated with cubic graphs, so satisfying
color-kinematics duality implicitly requires expressing the integrand
in terms of such graphs. As mentioned earlier, this is possible to do
even for a theory with quartic (or higher multiplicity) vertices by
multiplying and dividing by propagators.  In the case of the 4-point
2-loop example promised above, there are 14 cubic graph topologies
(shown in \cref{fig:MaxCuts}) ignoring tadpoles and bubble on external
leg (BEL) graphs.

With the integrand expressed in terms of cubic graphs, it is possible
to discuss the main ingredient in color-kinematics duality -- the
Jacobi relations.  The color factors in an integrand obey a set of
Jacobi identities.  Color-kinematics duality states that there
\emph{exists} a way of writing the kinematic numerators of the
integrand so that they obey the same set of Jacobi identities.  The
Jacobi identities can be divided into two categories: \emph{defining}
relations and \emph{boundary} relations.  Defining Jacobi relations
can be used to express any graph in terms of a basis.  For example,
the double box, crossed box, and penta-triangle are related to each
other as follows \be % [[11]]
\label{eq:TwoLoopJacobiOne}
\JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle =0\,,
\end{equation}
where the shared propagator is shown in red.  Using this relation the
crossed box can be written (or \emph{defined}) in terms of the other
two.  In this way, all of the 14 cubic 4-point 2-loop topologies can be
written in terms of a basis of the double box and penta-triangle.
(Any two of the double box, crossed box, and penta-triangle would
suffice for a basis but the planar basis is chosen for simplicity in
this work.)  The numerator of each of the two basis graphs receives
its own ansatz and the numerator of any other graph can be expressed
in terms of these two.
